---
title: "Project 2: Global Non-Linear Effect of Temperature on Economic Production"
author: "Andy Hultgren, Chester Harvey, and Rafael Grillo Avila"
date: "September 10, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Write-Up

#### Objective

The objective of our assignment was to recreate Figure 2, Panel A of Burke, Hsiang, and Miguel (2015), looking at the global non-linear (quadratic) relationship between annual average temperature andchange in log gross domestic product (GDP) per capita.  

Instead of plotting the change in log GDP (roughly measuring the change in growth rates of GDP), we wanted to measure the changes in GDP levels.

### Progress & Stumbling Blocks

We were able to recreate the data using the .do files, substituting GDP levels for log-GDP (reflecting growth rate of GDP).  

Three models were built, as described below.  The issues associated with Model 1 are also described, as well as the attempts to address one of those issues with Models 2 and 3.

#### Model 1 Format & Issues  

Model 1 (extension "_ds421_levels") regresses the level of GDP on a quadratic in country temperature, instead of growth in GDP on a quadratic in country temperature. Country-specific quadratic time controls are still used, just as a default, though it is clear that a country quadratic time control in a levels model is not controlling for the same variation in GDP as the same time trend in a growth model.

Now, this model has many problems associated with it that I think make it a bad idea to run. The first is that economic shocks in general affect a percentage of a country's GDP rather than its level of GDP. Think of some shock (e.g. a trade war, or anything): do we think it would cause a loss of e.g. $1,500 per capita whether you live in China, or Brazil, or the EU, or the US? No. Rather, we typically think of a loss of GDP of e.g. 5% as something that could be comparable across these places. (And data on GDP evolution support this intuition.) Two problems then fall out of this issue:

- *Problem 1:* GDP shocks should enter multiplicatively in the model rather than additively. i.e. the data generating process for country $i$ in time $t$ should be

(a) $GDP_{it} = some \thinspace shock_{it} \times GDP_{it-1}$

and the data generating process should not be

(b) $GDP_{it} = some \thinspace shock_{it} + GDP_{it-1}$

However model (b) is the model that is assumed when doing a levels regression. If you want to assume model (a), which I believe is the correct option, then you have to use the log model that the original authors use.

- *Problem 2:* Both models (a) and (b) above imply persistence of past shocks into the future. This introduces statistical concerns in the estimation process that also make a simple levels model (Model 1 above) a bad idea to run.


#### What do do about all this:

Both of these issues (Problem 1 and Problem 2) would, I think, be typical justifications for running a growth regression as the original authors did (and as Dell Jones and Olken did as well), and not a simple levels regression. But, we want to test the levels regression, so we are stuck with Problem 1.  For Problem 2, however, we can try first differencing to attempt to cancel out the persistent, past shocks. The next model does this:

- Model 2 (extension "_ds421_firstDiff"):  This model assumes data generating process (b) above, but takes first differences. Effectively:  $GDP_{it} - GDP_{it-1} = some \thinspace shock_{it} - some \thinspace shock_{it-1}$.  Importantly (for statistical purposes), $GDP_{it-1}$ is no longer on the right hand side of the regression model. We could run a battery of tests to ensure that first differencing fixes the persistent shock problem, but that would be too much work for this assignment.

One other note for this model: here the time controls are country fixed effect and a linear country time trend. In a first differences model, this is essentially similar to modeling a quadratic country time trend for the level (non-differenced) model. So, I make that change for consistency with Model 1 above.

- Model 3 (extension "_ds421_firstDiff_quadT"):  This tests the exact same model as Model 2, but includes quadratic country time trends, just to see the sensitivity to trends. Really none of these time controls is similar to having country quadratic time controls for the growth model, so I include Model 3 for the sake of being thorough.

Regarding the country time trends, we would have to do a lot more testing to actually determine what time controls are best given the data, and that is outside the scope of this assignment. Also, it might be futile given Problem 1 anyway.

### Results

Results are included alongside the code below, and in the GitHub file that this Markdown document is associated with.

## Code & Results

### Setup

Download necessary libraries and set up Figure 2A frame

```{r Environment}
require(maptools)
require(fields)
require(classInt)
require(plotrix)
require(dplyr)

# Basic layout

rm(list = ls())

#define a function for easy string pasting
"%&%" <- function(x, y) {
  paste(x,y,sep="")
} 

pdf(file="figures/MainFigs_Input/Figure2.pdf",width=10,height=5.5,useDingbats=F)

mat <- matrix(c(1,1,2,3,1,1,4,5),nrow=2,byrow=T)
layout(mat)
ax = 1.5  #scaling for axes
par(mar=c(4,4,2,1))
```

### Get Figure 2A Data

Get the data to make Figure 2, using outputs of .do files provided by Andy Hultgren, and focusing on Model 2 (first differences) specifically.

```{r Download Figure 2A Data}

# Model 2 Data

resp_diff <- read.csv("data/output/estimatedGlobalResponse_ds421_firstDiff.csv")
dta_diff <- read.csv("data/output/mainDataset.csv")
coef_diff <- read.csv("data/output/estimatedCoefficients_ds421_firstDiff.csv")
smpl_diff <- is.na(dta_diff$growthWDI) == F & is.na(dta_diff$UDel_temp_popweight) == F

# Original Data for comparison

resp <- read.csv("data/output/estimatedGlobalResponse.csv")
dta <- read.csv("data/output/mainDataset.csv")
smpl <- is.na(dta$growthWDI)==F & is.na(dta$UDel_temp_popweight)==F   #main estimation sample
coef <- read.csv("data/output/estimatedCoefficients.csv")
```
Edit data

```{r Edit Data}
# center response at optimum
x = resp_diff$x
mx = max(resp_diff$estimate)
est = resp_diff$estimate - mx
min90 = resp_diff$min90 - mx
max90 = resp_diff$max90 - mx

# Countries of interest
ctys = c('USA','CHN','DEU','JPN','IND','NGA','IDN','BRA','FRA','GBR')
ctt = c('US','CHN',"GER","JPN",'IND','NGA','INDO','BRA','FRA','UK')
``` 

```{r Plot Figure 2A}
#initialize plot, accounting for a change in levels for y-axis
plot(1,xlim=c(-2,30),ylim=c(min(min90), max(max90)),type="n",las=1,cex.axis=0.9)

# add vertical average temperature lines for selected countries
for (j in 1:length(ctys)) {
  tt = mean(dta_diff$UDel_temp_popweight[dta_diff$iso==ctys[j]],na.rm=T)
  segments(tt,-0.23,tt,0.15,lwd=0.5)
}

# plot CI and main effect
polygon(c(x,rev(x)),c(min90,rev(max90)),col="lightblue",border=NA)
lines(x,est,lwd=2)

# now add histograms at bottom
# first calculate percent of population and global gdp produced at each temperature bin, for our estimation sample
bins = seq(-7,30,0.5)
histtemp = dta_diff$UDel_temp_popweight[smpl]
histpop = dta_diff$Pop[smpl]
histgdp = dta_diff$TotGDP[smpl]
pop = gdp = c()
for (j in 1:(length(bins)-1)) {
  lo = bins[j]
  hi = bins[j+1]
  pop = c(pop,sum(histpop[histtemp>=lo & histtemp<hi]))
  gdp = c(gdp,sum(histgdp[histtemp>=lo & histtemp<hi]))
}
pop = pop/sum(pop)
gdp = gdp/sum(gdp)

#parameters that set where histograms go
dis = 0.055
base = -0.3

# now make histograms
#temperature
zz <- hist(histtemp,plot=F,breaks=bins)
cts = zz$counts/max(zz$counts)*0.05  #sets the height of the tallest bar to 0.05
rect(bins,base,bins+0.5,base+cts,col="red")
# pop
cts = pop/max(pop)*0.05
rect(bins,base-dis*(1),bins+0.5,base-dis*(1)+cts,col="grey")
# gdp
cts = gdp/max(gdp)*0.05
rect(bins,base-dis*(2),bins+0.5,base-dis*(2)+cts,col="black")

dev.off()
```

## Supplementary Material

### Readme Text

Readme for replicating Burke, Hsiang, and Miguel 2015

The data and code in the zipped directory BurkeHsiangMiguel2015_Replication.zip replicate all figures, tables, and estimates in Burke, Hsiang, Miguel 2015 (with the exception of Figure 1, which is constructed from figures in the literature).  

Full replication requires both R and Stata.  User-written programs for both R and Stata are required at various points, as highlighted in script headers where relevant.  Code was developed/implemented in Stata/MP 14, and R 3.2.0. 

Users of these data should cite Burke, Hsiang, Miguel 2015.  If you find a consequential error, or alternatively simply wish to congratulate us on a job well done, please email Marshall at mburke@stanford.edu.

Info on folders contained in replication data
—data/input, where input data is stored
—data/output, where generated data is stored that is then used to make figures.  
—figures/MainFigs_Input  where the panels for the figures in the main text are stored
—figures/ExtendedDataFigs_Input  where the panels for the figures in the extended data are stored
—script/  where scripts are stored

The main data file for the regressions is provided in both Stata 14 format (.dta), Stata 13 format (_Stata13.dta) as well as .csv. When importing the .csv into Stata, the option “case(preserve)” is needed to maintain consistency with the provided .do files.  

Instructions for replicating all figures/tables in main text and Extended Data
For all Stata and R code, user needs to set working directory to the directory where the files were unzipped.

1. First run all the Stata do files in the script/ folder :
	a. GenerateFigure2Data.do
	b. GenerateBoostrapData.do
	c. MakeExtendedDataFigure1.do, MakeExtendedDataFigure2.do, MakeExtendedDataFigure3.do, MakeExtendedDataTable1.do, MakeExtendedDataTable2.do

2.  Then run the R scripts generating the impact projections
	a. getTemperatureChange.R
	b. ComputeMainProjections.R
	c. ComputeDamageFunction.R

3.  Then run the desired MakeFigure or MakeExtendedDataFigure R script.


Instructions for generating specific figures in the main text
For all Stata and R code, user needs to set working directory to the directory where the files were unzipped.

To generate Figure 2, the user needs to run
1.  GenerateFigure2Data.do
2.  MakeFigure2.R

To generate Figure 3 or 4, the user needs to run:
3.  GenerateBootstrapData.do
4.  GetTemperatureChange.R  
5.  ComputeMainProjections.R
6.  And then run the desired MakeFigure3.R or MakeFigure4.R

To generate Figure 5, the user needs to run 3-5 above, and then run:
7. ComputeDamageFunction.R
8.  MakeFigure5.R